An FMI-Based Initialization Plugin for INTO-CPS Maestro 2

### 2020-06-29 background.tex
Updated some broken references

\todo{Christian should I add some pseudo code to explain?}
\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall \p \in LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

>>>>

\todo{Christian should I add some pseudo code to explain?}
\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall p \in LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

### 2020-06-29 realization.tex
Updated some broken references

\section{Realization of a Maestro 2 Plugin}\label{sc:implementation}
This approach has been realized as a Maestro 2 plugin that generates the \textit{Initialization}-phase of a co-simulation specification expressed in MaBL. It calculates the specification based on a list of FMU-components of a co-simulation scenario, and a specific plugin-configuration to let the user supply initial values of FMU variables with $causality=parameter$. A correction initialization specification can be calculated by the plugin if the co-simulation scenario is adhering to definition \ref{def:convergence}.

The plugin optimizes the initialization order by grouping operations that can be executed in \textit{parallel} to take advantage of FMI's ability to set or get multiple variables in bulk from the same FMU. The criteria for this optimization is defined in definition \ref{def:optimization}. 
The developed plugin has been tested on numerous co-simulation examples from the INTO-CPS universe\cite{Maestro} and compared with the existing \textit{Initializer} of Maestro. The plugin has been tested in the whole Maestro 2 pipeline to practically verify that it can be used in a co-simulation where the generated initialization procedure is integrated as a part of a complete co-simulation to verify the initialization procedure's integration with a master algorithm.

\subsection{Realization of the Topological Sorting}
The topological sorting algorithm (Tarjan's Algorithm) was implemented in Scala. The motivation for choosing Scala is its relation to JVM and the connection to Slang and the Logika framework\cite{inbook}. The connection to Slang and Logika will be investigated in the future. The Logika framework will be used to formally verify the implementation plugin and be used to explore how the contract-based nature of Slang can be used to obtain more reliable results of co-simulations. The topological sorting returns a topological order of strongly connected components. This order is the order of initialization, where the elements containing more than a single element denotes an algebraic loop that requires a particular initialization strategy.

\subsection{Verification of the approach}
The realized plugin and the developed approach have been verified in several different ways. Gomes et al. have in \cite{gomes_lucio_vangheluwe_2019} proved the correctness of using the topological order of a dependency graph of the interconnected variables to generate not only the initialization order but also the order for a co-simulation master-algorithm. The purpose of the publication by Gomes et al. is a little different from the one of this paper. Gomes et al. are interested in both generating an initialization and master-algorithm, meaning they also need to focus on the time of each FMU. They also define their graph based on FMU-operations ($Set, Get, doStep$) instead of interconnected variables, which is the approach described in this paper. This simplification is valid since this approach only cares about initialization and, therefore, does not need to care about the time of each FMU. This does that we can omit all the $doStep$ nodes and all edges connected to them from the graph by Gomes et al., this approach does, therefore, end up with a subgraph of the graph by Gomes et al. The proof from \cite{gomes_lucio_vangheluwe_2019} can therefore trivial be modified to the graph/approach presented in this paper.

>>>>

\section{Realization of a Maestro 2 Plugin}\label{sc:implementation}
This approach has been realized as a Maestro 2 plugin that generates the \textit{Initialization}-phase of a co-simulation specification expressed in MaBL. It calculates the specification based on a list of FMU-components of a co-simulation scenario, and a specific plugin-configuration to let the user supply initial values of FMU variables with $causality=parameter$. A correction initialization specification can be calculated by the plugin if the co-simulation scenario is adhering to the behavior dictated by definition given in \ref{def:convergence}.

The plugin optimizes the initialization order by grouping operations that can be executed in \textit{parallel} to take advantage of FMI's ability to \textit{set} or \textit{get} multiple variables in bulk from the same FMU. The criteria for this optimization is defined in definition \ref{def:optimization}. 
The developed plugin has been tested on numerous co-simulation examples from the INTO-CPS universe\cite{Maestro} and compared with the existing \textit{Initializer} of Maestro. The plugin has been tested in the whole Maestro 2 pipeline to practically verify that it can be used in a co-simulation where the generated initialization procedure is integrated as a part of a complete co-simulation to verify the initialization procedure's integration with a master algorithm.

\subsection{Realization of the Topological Sorting}
The topological sorting algorithm (Tarjan's Algorithm) was implemented in Scala. The motivation for choosing Scala is its relation to JVM and the connection to Slang and the Logika framework\cite{inbook}. The connection to Slang and Logika will be investigated in the future. The Logika framework will be used to formally verify the implemented plugin and be used to explore how the contract-based nature of Slang can be used to obtain more reliable results of co-simulations. The topological sorting returns a topological order of strongly connected components. This order is the order of initialization, where the elements containing more than a single element denotes an algebraic loop that requires a particular initialization strategy.

\subsection{Verification of the calculated Initialization Order}
The realized plugin has been verified in several different ways. Gomes et al. have in \cite{gomes_lucio_vangheluwe_2019} proved the correctness of using the topological order of a dependency graph of the interconnected variables to generate not only the initialization order but also the order of operations for a co-simulation step in master-algorithm. The purpose of the publication by Gomes et al. is a little different from the one of this paper. Gomes et al. are interested in both generating an initialization and master-algorithm, meaning they also need to focus on the time of each FMU. They also define their graph based on FMU-operations (\textit{Set, Get, doStep}) instead of interconnected variables, which is the approach described in this paper. This simplification is valid since this approach only cares about initialization and, therefore, does not need to care about the time of each FMU. This does that we can omit all the \textit{doStep} nodes and all edges connected to them from the graph by Gomes et al., this approach does, therefore, end up with a subgraph of the graph by Gomes et al. The proof from \cite{gomes_lucio_vangheluwe_2019} can therefore trivial be modified to the graph/approach presented in this paper.

### 2020-06-29 relatedwork.tex
Updated some broken references

Amalio et al. \cite{Amalio2016CheckingCo-simulation} has been investigating how to avoid algebraic loops in FMU based co-simulation scenarios by statically checking the architectural design of a CPS. The publication's purpose is like ours to avoid invalid co-simulation scenarios. Nevertheless, they achieve this by excluding scenarios containing algebraic loops. Their method is like ours also implemented in the real tool INTO-SysML, and it is based on using different kinds of formal methods (Theorem Proving and Model-checking) and will be an inspiration for the future work of formally verifying the plugin.

>>>>

Amalio et al. \cite{Amalio2016CheckingCo-simulation} has been investigating how to avoid algebraic loops in FMU based co-simulation scenarios by statically checking the architectural design of a CPS. The publication's purpose is like ours to avoid invalid co-simulation scenarios. Nevertheless, they achieve this by excluding scenarios containing algebraic loops. Their method is like ours also implemented in the real tool INTO-SysML\cite{Miyazawa2016INtegratedModelling}, and it is based on using different kinds of formal methods (Theorem Proving and Model-checking) and will be an inspiration for the future work of formally verifying the plugin.
